Highly Available Transactions: Virtues and Limitations

### 2013-04-28 writeup/conclusion.tex
some graphs and joe comments

\section{Conclusion}
The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. Accordingly, we expose a broad design space of Highly
Available Transactions (HATs), which can offer the key benefits of
highly available distributed systems---''always on'' operation during
partitions and low-latency operations---while also providing a family
of transactional isolation levels and replica semantics that have been
adopted in practice.  In taxonomizing this design space, we show that
many transactional guarantees are achievable with high availability,
while a few, like preventing Lost Update and Write Skew, are not. This
offers a long-overdue connection between storage semantics studied in
the database and distributed systems communities. The subsequent
challenge is to return to issues of performance and systems
architecture within the HAT design space.  Based on our understanding
of what is desired in the field and newfound knowledge of what is
possible to achieve, we believe HATs will be an area of great
opportunity in the coming years.

>>>>

\section{Conclusion}
The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. Accordingly, we expose a broad design space of Highly
Available Transactions (HATs), which can offer the key benefits of
highly available distributed systems---``always on'' operation during
partitions and low-latency operations---while also providing a family
of transactional isolation levels and replica semantics that have been
adopted in practice.  In taxonomizing this design space, we show that
many transactional guarantees are achievable with high availability,
while a few, like preventing Lost Update and Write Skew, are not. This
offers a useful connection between storage semantics studied in the
database and distributed systems communities. The next challenge is to
return to issues of performance and systems architecture within the
HAT design space.  Based on our understanding of what is desired in
the field and newfound knowledge of what is possible to achieve, we
believe HATs provide a large and useful design space to pursue.

### 2013-04-28 writeup/evaluation.tex
some graphs and joe comments

\vspace{.5em}\noindent\textbf{Serializability} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} requires at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ may require $T$ \texttt{lock}
operations and will require at least one \texttt{lock} and
\texttt{unlock} operation.  In a distributed environment, each of
these operations requires coordination, either with other database
servers or with a lock service. If this coordination mechanism is
unavailable, transactions cannot safely commit. Similarly, optimistic
concurrency control requires coordinating via a validation step, while
deterministic transaction scheduling~\cite{deterministic-scheduling}
requires a scheduler. Serializability under multi-version concurrency
control requires checking for update conflicts. All told, the reliance
on a globally agreed total order necessitates a minimum of one
round-trip to a designated master or coordination service for each of
these classic algorithms.  The cost of round trips will be determined
by the deployment environment, as we saw in
Section~\ref{sec:motivation}; we will demonstrate this cost on public
cloud infrastructure in Section~\ref{sec:prototype}.

@@

TPC-C consists of five transactions, capturing the operation of a
wholesale warehouse, including sales, payments, and deliveries. Two
transactions---\textit{Order-Status} and \textit{Stock-Level}---are
read-only and can be executed safely with HATs. Clients may read stale
data, but this does not violate TPC-C requirements and clients will
read their writes if they are sticky-available. Another transaction
type, \textit{Payment}, updates running balances for warehouses,
districts, and customer records and provides an audit trail. The
transaction is increment- and append-only, so all balance increase
operations commute, and TA allows the maintenance of foreign-key
integrity constraints (e.g., via \texttt{UPDATE/DELETE CASCADE}).

While three out of five transactions are easily achievable with
HATs, the remaining two transactions---\textit{New-Order} and
\textit{Delivery}---are not as simple. The New-Order transaction
places an order for a variable quantity of data items, updating
warehouse stock as needed. It selects a sales district, assigns the
order an ID number, adjusts the remaining warehouse stock, and writes
a placeholder entry for the pending order. The Delivery transaction
represents the fulfillment of a New-Order: it deletes the order from
the pending list, updates the customer's balance, updates the order's
carrier ID and delivery time, and updates the customer balance.

The New-Order transaction presents two challenges: ID assignment and
stock maintenance. First, each New-Order transaction requires a unique
ID number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can stil maintain uniqueness constraints. Second, the New-Order
transaction decrements inventory counts: what if the count becomes
negative?  Fortunately, TPC-C New-Order restocks each item's inventory
count (increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with more stock than in a
non-HAT-compliant implementation.

The Delivery transaction is challenging due to non-monotonicity. Each
Delivery deletes a pending order from the New-Order table and should
be idempotent in order to avoid billing a customer twice; this implies
a need to prevent Lost Update. This issue can be avoided by moving the
non-monotonicity to the real world---the carrier that picks up the
package for an order can ensure that she is the only carrier who has
done so---but cannot provide a correct execution with HATs
alone. However, according to distributed transaction
architects~\cite{entitygroup}, these compensatory actions are
relatively common in real-world business processes.

Throughout execution, TPC-C also requires the maintenance of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update rows in both tables atomically
via TA, can be easily maintained. Consistency Conditions 4 through 12
(3.3.2.4-12) can similarly be satisfied by applying updates atomically
across tables. Consistency Conditions 2 and 3 (3.3.2.2-3) concern
order ID assignment and are problematic. Finally, while TPC-C is not
subject to multi-key anomalies, we note that many TPC-E isolation
tests (i.e., simultaneously modifying a product description and its
thumbnail) are also achievable using HATs.

In summary, many---but not all---TPC-C transactions are well-served by
HATs. The two problematic transactions---New-Order and Payment---rely
on non-monotonic state update. The former can be modified to ensure ID
uniqueness but not sequential ID ordering, while the latter is an
inherently non-monotonic action requiring external compensation or
stronger consistency protocols. Based on these experiences and
discussions with practitioners, we expect that, especially for
read-dominated workloads found in many online services, HAT guarantees
will provide useful semantics for many applications.

@@

\begin{figure}[t!]
\begin{center}
\hspace{2em}\includegraphics[width=.8\columnwidth]{figs/strategylegend.pdf}
\end{center}\vspace{-2em}
\begin{center}\textbf{Within \texttt{us-east}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/lan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=0.90\columnwidth]{figs/lan-threads-thru.pdf}
\begin{center}\textbf{Between \texttt{us-east} and \texttt{us-west-2}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/wan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=0.90\columnwidth]{figs/wan-threads-thru.pdf}
\caption{YCSB performance for two clusters of three servers each
  deployed within a single datacenter and cross-datacenters.}
\label{fig:wan-exp}
\end{figure}

>>>>

\vspace{.5em}\noindent\textbf{Serializability} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} require at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ may require $T$ \texttt{lock}
operations and will require at least one \texttt{lock} and one
\texttt{unlock} operation.  In a distributed environment, each of
these operations requires coordination, either with other database
servers or with a lock service. If this coordination mechanism is
unavailable, transactions cannot safely commit. Similarly, optimistic
concurrency control requires coordinating via a validation step, while
deterministic transaction scheduling~\cite{deterministic-scheduling}
requires a scheduler. Serializability under multi-version concurrency
control requires checking for update conflicts. All told, the reliance
on a globally agreed total order necessitates a minimum of one
round-trip to a designated master or coordination service for each of
these classic algorithms.  The cost of round trips will be determined
by the deployment environment, as we saw in
Section~\ref{sec:motivation}; we will demonstrate this cost on public
cloud infrastructure in Section~\ref{sec:prototype}.

@@

TPC-C consists of five transactions, capturing the operation of a
wholesale warehouse, including sales, payments, and deliveries. Two
transactions---\textit{Order-Status} and \textit{Stock-Level}---are
read-only and can be executed safely with HATs. Clients may read stale
data, but this does not violate TPC-C requirements and clients will
read their writes if they are sticky-available. Another transaction
type, \textit{Payment}, updates running balances for warehouses,
districts, and customer records and provides an audit trail. The
transaction is monotonic---increment- and append-only---so all balance
increase operations commute, and TA allows the maintenance of
foreign-key integrity constraints (e.g., via \texttt{UPDATE/DELETE
  CASCADE}).

\vspace{.5em}\noindent\textit{New-Order and Delivery.} While three out of
five transactions are easily achievable with HATs, the remaining two
transactions are not as simple. The New-Order transaction places an
order for a variable quantity of data items, updating warehouse stock
as needed. It selects a sales district, assigns the order an ID
number, adjusts the remaining warehouse stock, and writes a
placeholder entry for the pending order. The Delivery transaction
represents the fulfillment of a New-Order: it deletes the order from
the pending list, updates the customer's balance, updates the order's
carrier ID and delivery time, and updates the customer balance.

\vspace{.5em}\noindent\textit{IDs and decrements.} The New-Order transaction presents two challenges: ID assignment and
stock maintenance. First, each New-Order transaction requires a unique
ID number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can still maintain uniqueness constraints. Second, the New-Order
transaction decrements inventory counts: what if the count becomes
negative?  Fortunately, TPC-C New-Order restocks each item's inventory
count (increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with more stock than in a
non-HAT-compliant implementation.

\vspace{.5em}\noindent\textit{TPC-C Non-monotonicity.} The Delivery
transaction is challenging due to non-monotonicity. Each Delivery
deletes a pending order from the New-Order table and should be
idempotent in order to avoid billing a customer twice; this implies a
need to prevent Lost Update. This issue can be avoided by moving the
non-monotonicity to the real world---the carrier that picks up the
package for an order can ensure that no other carrier will do so---but
cannot provide a correct execution with HATs alone. However, according
to distributed transaction architects~\cite{entitygroup}, these
compensatory actions are relatively common in real-world business
processes.

\vspace{.5em}\noindent\textit{Integrity Constraints.} Throughout execution, TPC-C also requires the maintenance of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update rows in both tables atomically
via TA, can be easily maintained. Consistency Conditions 4 through 12
(3.3.2.4-12) can similarly be satisfied by applying updates atomically
across tables. Consistency Conditions 2 and 3 (3.3.2.2-3) concern
order ID assignment and are problematic. Finally, while TPC-C is not
subject to multi-key anomalies, we note that many TPC-E isolation
tests (i.e., simultaneously modifying a product description and its
thumbnail) are also achievable using HATs.

\vspace{.5em}\noindent\textit{Summary.} Many---but not all---TPC-C transactions are well-served by
HATs. The two problematic transactions---New-Order and Payment---rely
on non-monotonic state update. The former can be modified to ensure ID
uniqueness but not sequential ID ordering, while the latter is an
inherently non-monotonic action requiring external compensation or
stronger consistency protocols. Based on these experiences and
discussions with practitioners, we expect that, especially for
read-dominated workloads found in many online services, HAT guarantees
will provide useful semantics for many applications.

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=.8\columnwidth]{figs/strategylegend.pdf}
\end{center}\vspace{-2em}
\begin{center}\small\textbf{Within \texttt{us-east (VA)}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/finals/2lan-threads-lats-log.pdf}\vspace{-.5em}
\includegraphics[width=0.90\columnwidth]{figs/finals/2lan-threads-thru.pdf}
\begin{center}\small\textbf{Between \texttt{us-east (CA)} and \texttt{us-west-2 (OR)}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/finals/2wan-threads-lats-log.pdf}\vspace{-.5em}
\includegraphics[width=0.90\columnwidth]{figs/finals/2wan-threads-thru.pdf}
\begin{center}\small Between \textbf{\texttt{us-east} \texttt{(VA)}, \texttt{us-west-1} \texttt{(CA)},\\ \texttt{us-west-2} \texttt{(OR)}, \texttt{eu-west} \texttt{(IR)}, \texttt{ap-northeast} \texttt{(SI)}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/finals/5wan-threads-lats-log.pdf}\vspace{--.5em}
\includegraphics[width=0.90\columnwidth]{figs/finals/5wan-threads-thru.pdf}
\caption{YCSB performance for two clusters of five servers each
  deployed within a single datacenter and cross-datacenters.}
\label{fig:wan-exp}
\end{figure}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.90\columnwidth]{figs/finals/scaleout-thru.pdf}
%\includegraphics[width=0.90\columnwidth]{figs/finals/scaleout-thru-perserver.pdf}
\end{center}\vspace{-2em}
\caption{Scale-out of TA, Eventual, and RC.}
\label{fig:scaleout}
\end{figure}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.90\columnwidth]{figs/finals/txnlen-thru.pdf}
\end{center}\vspace{-2em}
\caption{Transaction length versus throughput.}
\label{fig:txlen}
\end{figure}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.90\columnwidth]{figs/finals/wprop-thru.pdf}
\end{center}\vspace{-2em}
\caption{Proportion of reads and writes versus throughput.}
\label{fig:rprop}
\end{figure}

### 2013-04-28 writeup/futurework.tex
some graphs and joe comments

\textbf{What are the limits of HAT semantics?}  In this work, we have
taxonomized several well-defined and well-documented semantic
properties. This is useful as it contextualizes existing database and
distributed systems semantics but does not provide guidance as to what
consistency models are the ``strongest''---or admit the smallest set
of executions---within each availability class. A recent report from
UT Austin shows that no model stronger than causal consistency can be
achieved in a sticky highly available, one-way convergent
system~\cite{cac}, but there are infinite highly available and sticky
highly available models to consider. For example, always returning the
value $1$ in response to reads is incomparable with (i.e., neither
stronger nor weaker than) causal consistency: returning $1$ does not
respect causality, and causal consistency returns values other than
$1$. There are actually infintely many incomparable models (vacuously:
return $1$, return $2$, and so on). Are there \textit{meaningful}
properties that delineate the boundaries between availability levels?
Is there a ``meta-property'' that exactly characterizes what
properties are achievable with high availability or sticky high
availability? This is an area of ongoing research within the Berkeley
HAT project. \pbnote{SHORTEN}

@@

\textbf{Where do semantics-based concurrency control and weakened
  failure models lie in the HAT landscape?} The vast literature on
semantics-based concurrency control indicates when coordination can be
avoided while providing ``consistent'' operation for limited failure
scenarios. We have performed a narrow case study, but a general
taxonomization of techniques such as escrow
transactions~\cite{escrow}, Sagas~\cite{sagas}, and other long-running
transaction support could help address the challenge of programming in
a weakly consistent environment. HAT guarantees can also be
strengthened by considering weaker failure models such as networks
with bounded asynchrony. Additionally, while we have focused on
systems that provide a single availability mode, there are several
interesting hybrids to consider. For example, a system may provide
strong semantics in the absence of partitions, but, in the presence of
partitions, ``lose'' guarantees by having clients reconnect to an
available replica. This provides ``best effort'' guarantees but, to
provide availability, sacrifices them in the event of partitions.

>>>>

 A recent report from UT Austin shows that no model stronger than
 causal consistency can be achieved in a sticky highly available,
 one-way convergent system~\cite{cac}, but there are infinite highly
 available and sticky highly available models to consider. For
 example, always returning the value $1$ in response to reads is
 incomparable with (i.e., neither stronger nor weaker than) causal
 consistency: returning $1$ does not respect causality, and causal
 consistency returns values other than $1$. There are actually
 infintely many incomparable models (vacuously: return $1$, return
 $2$, and so on). \textbf{What are the limits of HAT semantics?}  Are
 there \textit{meaningful} properties that delineate the boundaries
 between availability levels?  Is there a ``meta-property'' that
 exactly characterizes what properties are achievable with high
 availability or sticky high availability?

@@

\textbf{Where do semantics-based concurrency control and weakened
  failure models lie in the HAT landscape?} The vast literature on
semantics-based concurrency control indicates when coordination can be
avoided while providing ``consistent'' operation for limited failure
scenarios. We have performed a narrow case study of TPC-C, but a
general taxonomization of techniques such as escrow
transactions~\cite{escrow}, Sagas~\cite{sagas}, and other long-running
transaction support could help address the challenge of programming in
a weakly consistent environment. HAT guarantees can also be
strengthened by considering weaker failure models such as networks
with bounded asynchrony. Additionally, while we have focused on
systems that provide a single availability mode, there are several
interesting hybrids to consider. For example, a system may provide
strong semantics in the absence of partitions, but, in the presence of
partitions, ``lose'' guarantees by having clients reconnect to an
available replica. This provides ``best effort'' guarantees but, to
provide availability, sacrifices them in the event of partitions.
